```{r setoptions, echo = TRUE}
# opts_chunk$set(echo = TRUE, results = "show", cache = TRUE)
```

# Reproducible Research: Peer Assessment 1

## Loading and preprocessing the data
### most calculations are performed here
```{r}
if(!file.exists("activity.zip")) stop("Cannot find archive activity.zip")
if(!file.exists("activity.csv")) unzip("activity.zip")
df <- read.csv("activity.csv", colClasses = c("integer", "Date", "integer"))  # unmodified
noNA <- df[which(!is.na(df$steps)),]                  # df with na removed
stepsDaily <- tapply(df$steps, df$date, sum)          # total steps per day
steps5min <- tapply(noNA$steps,noNA$interval,median)  # total steps per 5 minute interval
mean5min <- tapply(noNA$steps,noNA$interval,mean)     # mean steps per 5 minute interval
median5min <- tapply(noNA$steps,noNA$interval,median) # median steps per 5 minute interval
mean5minAll <- round(mean(mean5min, na.rm=TRUE))      # overall mean for 5 minute intervals

```

## What is mean total number of steps taken per day?
```{r}
plot(stepsDaily, type="h", xlab="day", ylab="number of steps")
print(paste(c("daily steps mean:", round(mean(stepsDaily, na.rm=TRUE)))))
print(paste(c("daily steps median:", round(median(stepsDaily, na.rm=TRUE)))))
```
## What is the average daily activity pattern?
```{r}
plot(mean5min, type="l", xlab="5 minute intervals", ylab="Number of steps (mean)")
print(paste(c("5 minute steps mean:", mean5minAll)))
print(paste(c("5 minute steps median:", round(median(median5min, na.rm=TRUE)))))
```

## Imputing missing values
Some rows have missing values, that is, some rows have NA as the value for Steps.
These are set to the mean value for all 5 minute periods by , for instance by setting those 
values to the overall mean value for 5-minute periods (i.e. 37).  
Here we create a second data frame, df2, which does this.
```{r}
dfNA <- is.na(df$steps)  # logical vector of NA rows
df2 <- df
df2[which(dfNA), "steps"] <- mean5minAll   # set them to the overall 5 minute mean
df2stepsDaily <- tapply(df2$steps, df2$date, sum)     # total steps per day for second dataset
```
## Plot using missing data and calculating the mean and median steps:
```{r}
plot(df2stepsDaily, type="h", xlab="day", ylab="number of steps")
print(paste(c("daily steps mean:", round(mean(df2stepsDaily, na.rm=TRUE)))))
print(paste(c("daily steps median:", round(median(df2stepsDaily, na.rm=TRUE)))))
```
We see a slight decrease in both the mean and median steps per day, (10766 | 10765) and (10766 | 10656).


## Are there differences in activity patterns between weekdays and weekends?





